Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2022-05-18 README.Rmd
add problems
small changes

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
output:
  bookdown::pdf_document2:
    latex_engine: "xelatex"
    toc: false
linestretch: 2
abstract: |
  Preregistration is becoming increasingly popular in the psychological sciences.
  At the same time, many researchers struggle to make precise predictions and feel overwhelmed by the need to prespecify every detail of their analysis plan.
  Should they surrender and discard the idea of preregistration altogether?
  Not at all.
  We argue for the benefits of preregistration beyond strictly confirmatory studies.
  Grounded in Bayesian reasoning, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigour.
  This objective rests on the relevance of "theoretical risk", a generalization of type-I error rate, for judging the evidential support for theories.
  We propose that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
  We argue that this reformulation extends the utility of preregistration and is more closely aligned with researchers intuition.
  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took.
  This uncertainty severely limits the utility of the researchers' evidence.
---

@@

@@

Deciding if a study is confirmatory or exploratory is a judgment made by the researcher assessing it.
The issues above arise by delegating this judgment to a simple decision rule.
That is, if something is preregistered, it is confirmatory, and if it is not preregistered, it is exploratory.
This rule is often assumed implicitly as a rule of thumb.
Other times it is stated as the explicit goal of preregistration.
In both cases, it leads to confusion when applied outside specific bounds.
Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
To uphold the simplicity of the decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what.
Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.

We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk", which is inversely related to type I error rate.
We then propose two interpretations of how theoretical risk is influenced by preregistration.
The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently limiting type I error.
The second interpretation is our main contribution and shows the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes.
Under this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but the uncertainty in judging the theoretical risk is minimized.
This changed viewpoint is crucial for the following reason.
As the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favouring the theory.
As we will later show, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privilege to operate outside of these incentives.
We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.

@@

The first assumption leads to our adoption of a Bayesian framework.
Our rationale is as follows.
Researchers that decide to conduct a study are akin to choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument.
This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in choosing a study.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the famous Bayes formula, on which we will heavily rely for our arguments.
Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
Our conclusions can be applied to bayesian and frequentist methods alike.
Of course, frequentist methods are traditionally viewed from a Popperian philosophy of science, but our conclusions are compatible with this view and notable later advancements like the error statistical view.
However, those philosophies of science require a much richer set of assumptions that are not quite necessary for the points we want to make about the value of preregistration.

@@

$P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
In carnaps tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.
We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.

>>>>

---
title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
output:
  bookdown::pdf_document2:
    latex_engine: "xelatex"
    toc: false
linestretch: 1
abstract: |
  Preregistration is becoming increasingly popular in the psychological sciences.
  At the same time, many researchers struggle to make precise predictions and feel overwhelmed by the need to prespecify every detail of their analysis plan.
  Should they surrender and discard the idea of preregistration altogether?
  Not at all.
  We argue for the benefits of preregistration beyond strictly confirmatory studies.
  Grounded in Bayesian reasoning, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigour.
  This objective rests on the relevance of "theoretical risk", a generalization of type-I error rate, for judging the evidential support for theories.
  We propose that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
  We argue that this reformulation extends the utility of preregistration and is more closely aligned with researchers intuition.
  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took which severely limits the utility of the researchers' evidence.
  Preregistration effectively reduces this uncertainty and therefore generates more persuasiveness evidence.
---

@@

<!--problems to solve:
* I feel like I am describing a ruby goldberg machine here and expect the reader to understand what happens when you move the spoon
* what about falsificationists, every time I write a paragraph relating to I delete it and try to move it a discussion
* references!!!
-->

@@

Deciding if a study is confirmatory or exploratory is a judgment made by the researcher assessing it.
The issues above arise by delegating this judgment to a simple decision rule.
That is, if something is preregistered, it is confirmatory, and if it is not preregistered, it is exploratory.
This rule of thumb is often assumed implicitly.
Other times it is stated as the explicit goal of preregistration.
In both cases, it leads to confusion when applied outside specific bounds.
Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
To uphold the simplicity of the decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what.
Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.

We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" (a term borrowed from Meehl), which is inversely related to type I error rate.
We then propose two interpretations of how theoretical risk is influenced by preregistration.
The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently limiting type I error.
The second interpretation is our main contribution and shows the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes.
Under this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but the uncertainty in judging the theoretical risk is minimized.
This changed viewpoint is crucial for the following reason.
As the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favouring the theory.
As we will later show, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privilege to operate outside of these incentives.
We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.

@@

The first assumption leads to our adoption of a Bayesian framework.
Our rationale is as follows.
Researchers that decide to conduct a study are akin to choosing a study to bet on.
They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument.
This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in choosing a study.
Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
The axioms of probability are sufficient to derive the famous Bayes formula, on which we will heavily rely for our arguments.
Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.

@@

$P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
In carnaps tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.[^corroboration]
<!--I think I get myself in hot water here, one the one hand I claim compatibility of hypo-deductive methods and the other hand I appeal to carnap!!!-->
We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.

[^corroboration]: This move to assume measures of corroboration is a leap of faith and not covered by the Dutch book argument we make earlier, so a falsificationist can rightfully reject it.
